---
output: github_document
bibliography: references/References.bib
biblio-style: apalike
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  echo = TRUE, 
  warning = FALSE, 
  error = FALSE, 
  message = FALSE,
  fig.align = 'center',
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "70%",
  collapse = TRUE,
  comment = "#>",
  fig.path = "figures/",
  tidy = FALSE
)
```

# `sgrqr`  

### Scoring the St George's Respiratory Questionnaire


```{r sri, echo=FALSE, fig.align = 'left', out.width=200, fig.asp = 1}
logo <- "figures/lungblue.png"
sgrq <- hexSticker::sticker(logo, 
               package="sgrqr", 
               s_x = 1, # horizontal position of subplot
               s_y = 1.2, # vertical position of subplot
               s_width = 0.6, # width of subplot
               s_height = 0.6, # height of subplot
               p_size = 14, # font size
               p_x = 1, # horizontal position of font
               p_y = .53, # vertical position of font
               p_color = "#0EB4DE", # font colour
               h_size = 2, # hexagon border size
               h_fill = "#0E163D", # hexagon fill colour
               h_color = "#0EB4DE", # hexagon border colour
               url = "St Georg Respiratory Questionnaire (SGRQ)", # url at lower border
               u_color = "#0EB4DE",
               u_size = 3,
               spotlight = TRUE,
               l_y = 1.05, 
               l_x = 1.25,
               l_height = 2.5,
               l_width = 2.25)
sgrq
```


## Package Description

With `score_sgrq` the package contains a function for scoring the [*St George's Respiratory Questionnaire*](http://www.healthstatus.sgul.ac.uk/SGRQ_download/Original%20English%20version.pdf) (SGRQ). In addition, `sgrqr` provides two simulated data frames containing SRI items:

* `sgrq.full`: Data frame with no missing values;
* `sgrq.na`: Data frame with some missing values.

## Installation

```{r, eval=FALSE}
devtools::install_github("nrkoehler/sgrqr")
